---
title: England Premier 2016/17 Part One, loading data
author: Stephen MacDonald
date: '2017-11-20'
categories:
  - england
  - premier league
tags:
  - exploratory analysis
slug: england-premier-2016-17
---

This is the first post on this site actually dealing with looking at football data. The data used is freely available from football-data.co.uk. Having downloaded the data I will first load it into R and have the full data set loaded into my environment.The first few steps will be getting a feel for the data and what is contained within it.Before we proceed any further I will load the required packages from R into the session. The tidyverse package by Hadley Wickham is a work of art and will contain everything we need at this moment in time.

```{r warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
```
I then use the read_csv function from the readrpackge to load the dataset intomy session.Note: the data must be within the samedirectory asthe markdown document in orderfor it to be loaded in.

```{r warning=FALSE, error=FALSE, message=FALSE}
df.prem16_17 <- read_csv(file = "E0.csv")
```
Note I have named my new R object (containing the data for analysis) with the prefix df. This is with the assumption that the data is in fact in the form of a df. Using read_csv for our data loading ensures that this is the case - in fact it is a dataframe and a tibble - something that is provided by the tidy verse and we can discuss later. We can check what form our data is in with a simple command

```{r}
class(df.prem16_17)
```

```{r}
head(df.prem16_17)
glimpse(df.prem16_17)
```
